Easy as Pi: A Network Coding Raspberry Pi Testbed

### 2016-06-06 cross_compilation.tex
finished fabric

\subsection{Kodo python}

Often when simplicity and fast prototyping is of higher importance than speed
it is often desired to use a scripting language rather than the often faster
compiled language. Kodo-python provides roughly similar bindings to the
underlying C++11 implementation of Kodo as Kodo-cpp does. This means that
python calls to Kodo-python are actually executed in the underlying C++11
compilied code.

Parsing calls and variables between Python and C++ compiled code is
not free, but this small introduced overhead may in many cases be worth
it to take advantage of Pythons and other scripting languages many
benefits.

Below is a short procedure to use Kodo-python in a \ac{Raspi}.

One advantage of Python compared to Kodo-cpp is that only the Kodo-python
library needs to be cross compiled. The python scripts using the Kodo
library needs not to be cross compiled but simply copied to the same
directory as .

Kodo-python is available at \url{https://github.com/steinwurf/kodo-python}.

\begin{lstlisting}[]
$ apt-get update
$ apt-get install python build-essential libpython-dev python-dev
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
% IN ARCH LINUX, MOD_PYTHON NEEDS TO BE INSTALLED

Create a directory to hold our code:
\begin{lstlisting}[]
$ CODEDIR="${HOME}/code"
$ mkdir -p ${CODEDIR}
$ cd ${CODEDIR}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\begin{lstlisting}[]
$ git clone git@github.com:steinwurf/kodo-python.git 
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\begin{lstlisting}[]
$ cd kodo-python
$ python2.7 waf configure --cxx_mkspec=cxx_raspberry_gxx49_arm
$ python waf build
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

KODO-PYTHON IS NOT WORKING!! WAITING FOR HELP ON STEINWURF DEV.

WRITE HOW TO RUN A KODO PYTHON SCRIPT IN THE RASPIS

\subsection{fabric}

@@

%SHORT INTRO. EXPLAIN A FEW USECASES FOR FABRIC. E.G. REBOOT, INSTALL PACKAGE, RUN SCRIPT ON MULTIPLE RASPIS. COPY FILES TO/FROM PC AND MULTIPLE RASPIS

>>>>

%\subsection{Kodo python}
%
%Often when simplicity and fast prototyping is of higher importance than speed
%it is often desired to use a scripting language rather than the often faster
%compiled language. Kodo-python provides roughly similar bindings to the
%underlying C++11 implementation of Kodo as Kodo-cpp does. This means that
%python calls to Kodo-python are actually executed in the underlying C++11
%compilied code.
%
%Parsing calls and variables between Python and C++ compiled code is
%not free, but this small introduced overhead may in many cases be worth
%it to take advantage of Pythons and other scripting languages many
%benefits.
%
%Below is a short procedure to use Kodo-python in a \ac{Raspi}.
%
%One advantage of Python compared to Kodo-cpp is that only the Kodo-python
%library needs to be cross compiled. The python scripts using the Kodo
%library needs not to be cross compiled but simply copied to the same
%directory as .
%
%Kodo-python is available at \url{https://github.com/steinwurf/kodo-python}.
%
%
%
%\begin{lstlisting}[]
%$ apt-get update
%$ apt-get install python build-essential libpython-dev python-dev
%\end{lstlisting}
%\FloatBarrier
%\vspace{-5mm}
%  % IN ARCH LINUX, MOD_PYTHON NEEDS TO BE INSTALLED
%
%
%Create a directory to hold our code:
%\begin{lstlisting}[]
%$ CODEDIR="${HOME}/code"
%$ mkdir -p ${CODEDIR}
%$ cd ${CODEDIR}
%\end{lstlisting}
%\FloatBarrier
%\vspace{-5mm}
%
%\begin{lstlisting}[]
%$ git clone git@github.com:steinwurf/kodo-python.git 
%\end{lstlisting}
%\FloatBarrier
%\vspace{-5mm}
%
%\begin{lstlisting}[]
%$ cd kodo-python
%$ python2.7 waf configure --cxx_mkspec=cxx_raspberry_gxx49_arm
%$ python waf build
%\end{lstlisting}
%\FloatBarrier
%\vspace{-5mm}
%
%KODO-PYTHON IS NOT WORKING!! WAITING FOR HELP ON STEINWURF DEV.
%
%WRITE HOW TO RUN A KODO PYTHON SCRIPT IN THE RASPIS

\subsection{Fabric}

@@

### 2016-06-06 testbed.tex

\begin{figure}[ht!]
\centering
\includegraphics[width=0.6\textwidth]{images/testbed_setup2.eps}
\caption{Testbed setup}
\label{fig:testbed_setup}
\end{figure}

@@

\textbf{hashmark (\#) means to perform operation as root, dollar (\$) means
to perform operation as user}

@@

@@

% Root example
\begin{lstlisting}[]
$ whoami
chres
$ sudo whoami
root
\end{lstlisting}
\FloatBarrier

@@

After Raspbian lite has been unpacked, there should be an .img file in the directory.
"fdisk" can be used to display what is inside. Pass the arguments "-u sectors" to
display the sizes in sectors and "-l" to display partitions and exit.

@@

The image contains two partitions. The first one of starting at sector 8192 and
the other partition starting at sector 137216. The partitions is of type FAT32
and Linux and have sizes 63M and 1.2G. This indicates that the first partition
is most likely a boot partition and the second partition is a traditional file
system. Actually, it is a linux root filesystem (i.e. /).

%We see that the boot and root partitions starts at sector 8192 and 137216 respectivly.
%We also notice that each sector is 512 bytes.

%On newer systems, we can mount the device easier using "losetup --show -f -P IMAGE"

%Mount root partition:

>>>>

\begin{figure}[ht!]
\centering
\includegraphics[width=0.85\textwidth]{images/testbed_setup2.eps}
\caption{Testbed setup}
\label{fig:testbed_setup}
\end{figure}

@@

@@

Hashmark (\#) and dollar (\$) will be used throughout the procedure
to indicate if an operation should be run with root permissions or
user permissions.
%\textbf{hashmark (\#) means to perform operation as root, dollar (\$) means
%to perform operation as user}

@@

% Root example
\begin{lstlisting}[]
$ whoami
<USERNAME>
$ sudo whoami
root
\end{lstlisting}
\FloatBarrier

@@

\subsection{The image}

After Raspbian lite has been unpacked, there should be an .img file in the directory.
"fdisk" can be used to display what is inside. Pass the arguments "-u sectors" to
display the sizes in sectors and "-l" to display partitions and exit fdisk again.

@@

The image is in total 2709504 sectors (1.3GiB) in size and contains two
partitions. The first one of starting at sector 8192 and
the other partition starting at sector 137216. The partitions is of type FAT32
and Linux and have sizes 63M and 1.2G. This indicates that the first partition
is most likely a boot partition and the second partition is a traditional linux file
system. Actually, it is the linux root filesystem (i.e. /).

\subsection{(Optional) Resize image}
1.2G is not very much to hold the root filesystem given that it is desired to
in later steps alter the files and install more packages. It may therefore be
neccessary to increase the partition size slightly.
The following steps will therefore illustrate how image and root filesystem
can be expanded by one \ac{GB}.

First, expand the image one \ac{GB}:
\begin{lstlisting}[]
# dd if=/dev/zero bs=1M count=1024 >> ${IMAGE}.img
\end{lstlisting}
\FloatBarrier

Use \texttt{fdisk} to check that the image is now one \ac{GB} bigger:
\begin{lstlisting}[literate={DOLLAR}{\$}1]
DOLLAR fdisk -u sectors -l ${IMAGE}.img
Disk 2016-05-27-raspbian-jessie-lite.img: 2.3 GiB, 2461007872 bytes, 4806656 sectors
Units: sectors of 1 * 512 = 512 bytes
Sector size (logical/physical): 512 bytes / 512 bytes
I/O size (minimum/optimal): 512 bytes / 512 bytes
Disklabel type: dos
Disk identifier: 0x6fcf21f3

%We see that the boot and root partitions starts at sector 8192 and 137216 respectivly.
%We also notice that each sector is 512 bytes.
%On newer systems, we can mount the device easier using "losetup --show -f -P IMAGE"
%Mount root partition:
